---
title: "Agglometative Hierarchical Clustering for Water Chemistry by Site"
author: 'Margaret Hammond'
date: '2/16/2022'
output: 
  html_document:
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(janitor)
library(tidyverse)
library(palmerpenguins)
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```

### Summary
In this task, I will be analyzing stream water chemistry measurements taken in Santa Barbara area watersheds, beginning in 2000 (SBC LTER 2019). The stream water chemistry sites will be analyzed by agglomeration cluster analysis, and the results are shown in a dendrogram. This dendrogram shows which sites have similar water chemistry. 

### Wrangle the Data

```{r}
chem_sbr <- read_csv(here('data', 'sbc_lter_registered_stream_chemistry.csv')) %>% 
  clean_names() %>% 
  na_if(-999) %>% 
  na.omit() %>% 
  select(-timestamp_local) %>% 
  group_by(site_code) %>% 
  summarize(nh4_mean = mean(nh4_u_m),
            tdn_mean = mean(tdn_u_m),
            tdp_mean = mean(tdp_u_m),
            tpc_mean = mean(tpc_u_m),
            tpn_mean = mean(tpn_u_m),
            tpp_mean = mean(tpp_u_m),
            tss_mean = mean(tss_mgper_liter),
            no3_mean = mean(no3_u_m),
            po4_mean = mean(po4_u_m),
            spec_cond_mean = mean(spec_cond_u_spercm))

water_chem_scale <- chem_sbr %>% 
  select(-site_code) %>% 
  scale()

rownames(water_chem_scale) <- chem_sbr$site_code
  
```
### Agglomerative Cluster Analysis

```{r}
site_dist <- dist(water_chem_scale, method = 'euclidean') ### creates distance matrix


hc_complete <- hclust(site_dist, method = 'complete') ### create the clustering

ggdendrogram(hc_complete, rotate = TRUE) +
  theme_minimal() +
  scale_colour_manual(values = c('red', 'blue')) +
  labs(x = 'Site',
       y = 'Height',
       title = 'Santa Barbara Sites by Water Chemistry',
       caption = 'Figure 1. Santa Barbara sites as clustered by chemistry levels of water 
       (SBC LTER 2019).') 



dend <- as.dendrogram(hc_complete)
color_branches(dend, k = 2, col = c('red', 'blue'))
dend %>% set("branches_k_color", 
             value = c('red', 'blue'),
             k = 2) %>% 
  plot(main = "Santa Barbara Sites by Water Chemistry",
       horiz = FALSE)


```

### Dendrogram Takeaways
  - ON02 and GV01 are significantly similar to one another and most different than the other sites. 
  - 
  - 

### Citation
Citation: Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 
