---
title: "beepboop"
output: html_document
name: 'Margaret Hammond'
theme: 'yeti'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(janitor)
library(tidyverse)
library(ggfortify)
library(palmerpenguins)
library(patchwork)
library(paletteer)
```


A useful descriptive introductory summary (3 - 4 sentences) that helps the audience understand the data (include a citation as necessary) and what you’ll be exploring
All of your organized and well-annotated code (with warnings/messages suppressed) you wrote to wrangle data then run PCA, and to create a professional looking PCA biplot that appears (with a caption) in your knitted HTML (optional: also include a nicely formatted screeplot)
A brief summary (nicely formatted and professionally written bulletpoints are fine) highlighting some major takeaways from your PCA that can be gleaned from the biplot (and screeplot, if included). 



```{r}
nutrients <- read_csv(here('data', 'usda_nutrients.csv')) %>% 
  clean_names() %>% 
    filter(food_group %in% c('Dairy and Egg Products', 'Fast Foods', 'Nut and Seed Products', 'Snacks', 'Vegetable and Vegetable Products'))

nutrients_pca <- nutrients %>% 
  select(energy_kcal, protein_g, fat_g, carb_g) %>% 
  scale() %>% 
  prcomp()
```

## Biplot

```{r}
 
 
autoplot(nutrients_pca,
         data = nutrients,
         loadings = TRUE,
         colour = "food_group",
         loadings.label = TRUE,
         loadings.colour = 'black',
         loadings.label.colour = 'black',
         loadings.label.vjust = -0.5) +
  scale_color_manual(values = c('blue', 'purple', 'orange', 'red', 'green')) +
  scale_fill_manual(values = c()) +
  theme_minimal()
```

```{r}
autoplot(nutrients_pca,
         data = nutrients,
         loadings = TRUE,
         colour = 'food_group',
         loadings.label = TRUE,
         loadings.colour = 'black',
         loadings.label.colour = 'black',
         loadings.label.vjust = -0.5) +
  scale_color_paletteer_d('nord::aurora') +
  theme_minimal()
```

### Citation
 U.S. Department of Agriculture, Agricultural Research Service. FoodData Central, 2019. fdc.nal.usda.gov.
